The K2 Bright Star Survey I: Methodology and Data Release

### 2019-08-16 paper/ms.tex
colour-magnitude diagram

% describe saturation
The \kepler detector saturates at a magnitude of $K_p \sim 11.3$ in both long- (30\,min) and short (1\,min)-cadence data, since these both represent sums of 6\,s exposures \citep{Gilliland2010}. For objects brighter than this, excess electrons `bleed' into adjacent pixels in both directions along the column containing the star. Simple aperture photometry (SAP) -- adding all the flux contained in a window around the bleed column -- has recovered light curves with precisions close to the photon noise limit. Examples treated in the nominal {\it Kepler\/} mission are the prototype classical radial pulsator RR\,Lyr \citep[$V=7.2$;][]{Kolenberg2011}, the solar-like pulsators 16\,Cyg\,AB \citep[$V\approx 6$;][]{2012ApJ...748L..10M,2013MNRAS.433.1262W,2015ApJ...811L..37M} and $\theta\,$Cyg \citep[$V=4.48$;][]{Guzik2016}, and the massive eclipsing binary V380\,Cyg \citep[$V=5.68$;][]{tkachenko2014}. In the nominal \kepler mission SAP was only attempted for a few bright stars, and in \ktwo, the larger spacecraft motion significantly increased the size of the required apertures for SAP photometry of very saturated stars, while also making their systematics more difficult to deal with. While the second-version pixel-level-decorrelation (PLD) pipeline EVEREST~2.0 was able to correct systematics in saturated SAP photometry \citep{everest2}, this is not possible for the very brightest stars whose bleed columns may run to the edge of the detector. Furthermore, bandwidth constraints meant that pixel data were not downloaded for many bright targets in \ktwo.

@@

% describe halo algorithm
The `TV-min' halo method was first described by \citet{White2017} and applied to the Pleiades' Seven Sisters. It was also applied to Aldebaran with further developments by \citet{Farr2018}. We follow the Optimized-Weight Linear (OWL) photometry concept described by Hogg \&  Foreman-Mackey (2014, unpublished: preprint  \href{https://github.com/davidwhogg/OWL/}{\nolinkurl{github.com/davidwhogg/OWL/}}) in our assumptions. We assume that a star has a wide PSF sampled by many pixels with different sensitivities. This PSF varies at most to a small extent in time. The star moves around on the detector within a small region. We assume that our time series consists of many epochs sampled with a nearly even cadence. We do not have metadata describing the spacecraft motion, pixel gains, PSF variations or other noise processes.

@@

The additional constraint beyond the OWL axioms is that some pixels are saturated, so that Simple Aperture Photometry (SAP) is inadvisable. Instead the measurements are made using the unsaturated pixels at the wings of the broad and structured PSF, with counts $p_ij$ where pixels are indexed by $j$ and epoch by $i$. We construct a light curve as a linear combination of these time series with weights $w_j$, so that flux $f_i$ at epoch $i$ is a linear

@@

%  improvements
Unlike other methods for calibrating \kepler systematics, other than the value of $\delta$, no knowledge of the spacecraft motion or the behaviour of an ensemble of other stars is used to inform our algorithm. The signal and the noise are jointly estimated from the data. The method is a self-calibration that is independent of the details of the systematics it is calibrating, operating on the assumption that a single signal is present across many individual time series which otherwise are contaminated by noise.

@@

We calculated halo light curves of 36\,Oph and their CDPPs for $k \in \{1,2,3\}$, and $\delta \in [1,50]$ for long cadence and for various values of $\delta \in [1,2500]$ for short cadence data. The results are displayed in Figures~\ref{cdpps_lc} and~\ref{cdpps_sc}. We found that for long cadence data, $k=1$ (TV) and a lag $\delta=10$ provide the best CDPP. As this is one cadence shorter than the 12~cadence thruster firing period, we can understand the optimum as suppressing systematics on the same timescale as they occur. On the other hand, for short cadence data, performance at short lags is very poor but the method performs similarly for $k \in \{1,2\}$ with slow improvement with larger $\delta$, and performs very poorly for $k=3$ at all lags.

We accordingly use a lag $\delta=10$ for all long cadence light curves, and a lag $\delta=300$ for short cadence for consistency in timescale with the long cadence processing.

@@

\begin{figure*}
\plotone{lag_opt_both.pdf}
\caption{Behaviour of long cadence 6.5~hour CDPP (bottom) and $4 c/d$ systematics power (top) for the quiet dwarf 36~Ophiuchi as a function of lag parameter $\delta$. CDPP shows a minimum for L1 norm and $\delta=10$, i.e. for objective function $Q_{1,10}$, which is marked with a blue dashed vertical line. This does not correspond to an optimum in systematic power, which is slightly lower for smaller $\delta$. Nevertheless, we have chosen $\delta=10$ for the light curves in this catalog because of its improvement in overall CDPP as a measure of planet detection efficiency and overall light curve quality.}
\label{cdpps_lc}
\end{figure*}

@@

\begin{figure}
\plotone{halo_pdc_comparison.png}
\caption{Correlation diagram of the \texttt{lightkurve}-computed 6.5\,hr CDPP for \ktwo Campaign~6 stars in the magnitude range $11.5 < Kp < 12.5$, as processed with the PDC pipeline ($x$-axis) and TV-min pipeline ($y$-axis), both after correction and whitening with \textsc{k2sc}. The severe outliers where halo significantly outperforms PDC are shown by individual inspection to consist of stars for which there is contamination, or for which the SAP aperture assigned by the \kepler pipeline is significantly smaller than the PSF.}
\label{fig:halovspdc}
\end{figure}

@@

The full sample of stars for which halo apertures were obtained is listed in Table~\ref{table_all}. Following the successful pilot observations of the Pleiades B~stars in Campaign~4, we proposed halo photometry through dedicated K2 Guest Observer Programs from Campaign 6 onwards. Target selection was performed by cross-matching Hipparcos \citep{leeuwen07} with the K2 Ecliptic Plane Input Catalog \citep[EPIC,][]{huber16} and selecting all targets on silicon brighter than $Kp < 6$ on silicon. M~giants which pulsate with periods that are long compared to a K2 campaign were removed. We requested short-cadence observations for a small number of unevolved stars for which the expected timescales of oscillations cannot be sufficiently sampled with long-cadence data, such as for $\delta$~Sct stars whose maximum frequencies can exceed the long-cadence Nyquist limit.

@@

% short cadence availability 
Seven stars in Campaign~13 and one in Campaign~16 were assigned short-cadence halo apertures. For these targets we have provided both long- and short-cadence reductions. Following the analysis in Section~\ref{method} showing the insensitivity of short cadence CDPP to lags longer than $\sim 100$\,cad and to $k \in {1,2}$, and for consistency with long cadence, we have adopted a 300~epoch lag (i.e. $30 \times$ the long-cadence lag of 10) and the L1 TV objective function. With their many time samples, the short-cadence stars are computationally intractable for the Gaussian Process model in \textsc{k2sc} and we present otherwise uncalibrated halo light curves.

@@

Analyses for some of our sample have been previously published, we include their light curves in this data release: the Pleiades Seven Sisters \citep{White2017}, $\alpha$~Tau \citep[Aldebaran;][]{Farr2018}, $\iota$~Lib \citep{Buysschaert2018}, and $\epsilon$~Tau \citep{Arentoft2019}, as well as $\rho$~Leo, which was studied with halo pixels but without our objective functions \citep{Aerts2018}.

@@

% interesting EB - 98 Tau
The chemically-peculiar A0V star 98~Tau is of special interest for studies of surface inhomogeneity. We detected variability with a fundamental period of 1.74~d with twice as much power at the first harmonic ($P = 0.87$\,d), which is consistent with $\alpha^2$\,CVn chemical spot modulation from a rapidly-rotating star. This star also experiences a transit of depth 0.03, which for a 1.87\,\rsun typical A0V star implies an 0.3\,\rsun stellar mass companion. There are an unusually high number of background stars in the same photometric aperture as 98~Tau, and these were not all detected by \texttt{deathstar} and significantly contaminated the resulting lightcurve. As a result it was necessary to manually flag these objects using the `interact' mode of \texttt{lightkurve}, as displayed in Figure~\ref{fig:98tau}. The eclipse is deep enough to be seen by eye in the diffuse light of 98~Tau using this interactive display, and is not associated with any of the background stars.

@@

% halo method
There are directions for improvement of the halo method itself, and for applying it beyond \kepler/\ktwo and TESS. It remains to be seen how well the method of optimizing convex objective functions can deal with significantly varying PSFs, such as from ground-based observations. The rapidly varying and moving seeing-limited PSF couples to flat field errors as is the case with \kepler, and leads to severe short-timescale instrumental noise. Self-calibration by the halo method, or a similar method, may permit improvements in ground-based photometry. Likewise, there may be other convex objective functions, including linear combinations of currently-used objective functions, which offer superior performance, for example by using combinations of different lagged functions to suppress systematics occurring at different timescales. The remaining unexplored space of convex objective functions suitable may offer significant improvements on existing self-calibration techniques in high-cadence photometry and related problems in astronomy.

>>>>

% describe saturation
The \kepler detector saturates at a magnitude of $K_p \sim 11.3$ in both long- (30\,min) and short (1\,min)-cadence data, since these both represent sums of 6\,s exposures \citep{Gilliland2010}. For objects brighter than this, excess electrons `bleed' into adjacent pixels in both directions along the column containing the star. Simple aperture photometry (SAP) -- adding all the flux contained in a window around the bleed column -- has recovered light curves with precisions close to the photon noise limit. Examples treated in the nominal {\it Kepler\/} mission are the prototype classical radial pulsator RR\,Lyr \citep[$V=7.2$;][]{Kolenberg2011}, the solar-like pulsators 16\,Cyg\,AB \citep[$V\approx 6$;][]{2012ApJ...748L..10M,2013MNRAS.433.1262W,2015ApJ...811L..37M} and $\theta\,$Cyg \citep[$V=4.48$;][]{Guzik2016}, and the massive eclipsing binary V380\,Cyg \citep[$V=5.68$;][]{tkachenko2014}. In the nominal \kepler mission SAP was only attempted for a few bright stars, and in \ktwo, the larger-amplitude spacecraft motion significantly increased the size of the required apertures for SAP photometry of very saturated stars, while also making their instrumental systematics more difficult to deal with. While the second-version pixel-level-decorrelation (PLD) pipeline EVEREST~2.0 was able to correct systematics in saturated SAP photometry \citep{everest2}, this is not possible for the very brightest stars whose bleed columns may run to the edge of the detector. Furthermore, bandwidth constraints meant that pixel data were not downloaded for many bright targets in \ktwo.

@@

% describe halo algorithm
The `TV-min' halo method was first described by \citet{White2017} and applied to the Pleiades' Seven Sisters. It was also applied to Aldebaran with further developments by \citet{Farr2018}. In this Section we will discuss some improvements made to the halo method since those publications, and describe tests of the method using saturated and unsaturated targets.

We follow the Optimized-Weight Linear (OWL) photometry concept described by Hogg \&  Foreman-Mackey (2014, unpublished: preprint  \href{https://github.com/davidwhogg/OWL/}{\nolinkurl{github.com/davidwhogg/OWL/}}) in our assumptions. We assume that a star has a wide PSF sampled by many pixels with different sensitivities. This PSF varies at most to a small extent in time. The star moves around on the detector within a small region. We assume that our time series consists of many epochs sampled with a nearly even cadence. We do not have metadata describing the spacecraft motion, pixel gains, PSF variations or other noise processes.

@@

The additional constraint beyond the OWL axioms is that some pixels are saturated, so that Simple Aperture Photometry (SAP) is inadvisable. Instead the measurements are made using the unsaturated pixels at the wings of the broad and structured PSF, with counts $p_{ij}$ where pixels are indexed by $j$ and epoch by $i$. We construct a light curve as a linear combination of these time series with weights $w_j$, so that flux $f_i$ at epoch $i$ is

@@

%  improvements
Unlike other methods for calibrating \kepler systematics, other than the value of~$\delta$, no knowledge of the spacecraft motion or the behaviour of an ensemble of other stars is used to inform our algorithm. The signal and the noise are jointly estimated from the data. The method is both self-calibrating, and is independent of the details of the systematics it is calibrating, operating on the assumption that a single signal is present across many individual time series which otherwise are contaminated by noise.

@@

We calculated halo light curves of 36\,Oph and their CDPPs for $k \in \{1,2,3\}$, and $\delta \in [1,50]$ for long cadence and for various values of $\delta \in [1,2500]$ for short-cadence data. The results are displayed in Figures~\ref{cdpps_lc} and~\ref{cdpps_sc}. We found that for long-cadence data, $k=1$ (TV) and a lag $\delta=10$ provide the best CDPP. As this is one cadence shorter than the 12~cadence thruster firing period, we can understand the optimum as suppressing systematics on the same timescale as they occur. On the other hand, for short-cadence data, performance at short lags is very poor but the method performs similarly for $k \in \{1,2\}$ with slow improvement with larger $\delta$, and performs very poorly for $k=3$ at all lags.

We accordingly use a lag $\delta=10$ for all long-cadence light curves, and a lag $\delta=300$ for short cadence for consistency in timescale with the long-cadence processing.

@@

\begin{figure*}
\plotone{lag_opt_both.pdf}
\caption{Behaviour of long-cadence 6.5~hour CDPP (bottom) and $4 c/d$ systematics power (top) for the quiet dwarf 36~Ophiuchi as a function of lag parameter $\delta$. CDPP shows a minimum for L1 norm and $\delta=10$, i.e. for objective function $Q_{1,10}$, which is marked with a blue dashed vertical line. This does not correspond to an optimum in systematic power, which is slightly lower for smaller $\delta$. Nevertheless, we have chosen $\delta=10$ for the light curves in this catalog because of its improvement in overall CDPP as a measure of planet detection efficiency and overall light curve quality.}
\label{cdpps_lc}
\end{figure*}

@@

\begin{figure}
\plotone{halo_pdc_comparison.pdf}
\caption{Correlation diagram of the \texttt{lightkurve}-computed 6.5\,hr CDPP for \ktwo Campaign~6 stars in the magnitude range $11.5 < Kp < 12.5$, as processed with the PDC pipeline ($x$-axis) and TV-min pipeline ($y$-axis), both after correction and whitening with \textsc{k2sc}. The severe outliers where halo significantly outperforms PDC are shown by individual inspection to consist of stars for which there is contamination, or for which the SAP aperture assigned by the \kepler pipeline is significantly smaller than the PSF.}
\label{fig:halovspdc}
\end{figure}

@@

The full sample of stars for which halo apertures were obtained is listed in Table~\ref{table_all}. A $B$, $V$ color-magnitude diagram is displayed in Figure~\ref{cmd_halo}, omitting the very-red carbon star HR~3541, whose $B-V$ color is 3.23. Following the successful pilot observations of the Pleiades B~stars in Campaign~4, we proposed halo photometry through dedicated K2 Guest Observer Programs from Campaign 6 onwards. Target selection was performed by cross-matching Hipparcos \citep{leeuwen07} with the K2 Ecliptic Plane Input Catalog \citep[EPIC,][]{huber16} and selecting all targets on silicon brighter than $Kp < 6$ on silicon. M~giants which pulsate with periods that are long compared to a K2 campaign were removed. We requested short-cadence observations for a small number of unevolved stars for which the expected timescales of oscillations cannot be sufficiently sampled with long-cadence data, such as for $\delta$~Sct stars whose maximum frequencies can exceed the long-cadence Nyquist limit.

\begin{figure}
\plotone{cmd_halo.pdf}
\caption{$B$, $V$ color-magnitude diagram of the halo sample overlaid on a random subset of K2 stars with high-SNR \emph{Gaia} cross-matches, from the \href{https://gaia-kepler.fun}{\nolinkurl{gaia-kepler.fun}} sample, with $B$ and $V$ magnitudes drawn from the EPIC catalog \citep{huber16}. We omit the very-red carbon star HR~3541, whose $B-V$ color is 3.23. An interactive version of this plot is available as supplementary online material.}
\label{cmd_halo}
\end{figure}

@@

% short cadence availability 
Seven stars in Campaign~13 and one in Campaign~16 were assigned short-cadence halo apertures. For these targets we have provided both long- and short-cadence reductions. Following the analysis in Section~\ref{method} showing the insensitivity of short-cadence CDPP to lags longer than $\sim 100$\,cad and to $k \in {1,2}$, and for consistency with long cadence, we have adopted a 300~epoch lag (i.e. $30 \times$ the long-cadence lag of 10) and the L1 TV objective function. With their many time samples, the short-cadence stars are computationally intractable for the Gaussian Process model in \textsc{k2sc} and we present otherwise uncalibrated halo light curves.

@@

Analyses for some of our sample have been previously published, we include their light curves in this data release: the Pleiades' Seven Sisters \citep{White2017}, $\alpha$~Tau \citep[Aldebaran;][]{Farr2018}, $\iota$~Lib \citep{Buysschaert2018}, and $\epsilon$~Tau \citep[Ain;][]{Arentoft2019}, as well as $\rho$~Leo, which was studied with halo pixels but without our objective functions \citep{Aerts2018}.

@@

% interesting EB - 98 Tau
The chemically-peculiar A0V star 98~Tau is of special interest for studies of surface inhomogeneity. We detected variability with a fundamental period of 1.74~d with twice as much power at the first harmonic ($P = 0.87$\,d), which is consistent with $\alpha^2$\,CVn chemical spot modulation from a rapidly-rotating star. This star also experiences a V-shaped transit of depth 0.16, which for a 1.87\,\rsun typical A0V star implies an grazing eclipse by a stellar mass companion. There are an unusually high number of background stars in the same photometric aperture as 98~Tau, and these were not all detected by \texttt{deathstar} and significantly contaminated the resulting lightcurve. As a result it was necessary to manually flag these objects using the `interact' mode of \texttt{lightkurve}, as displayed in Figure~\ref{fig:98tau}. The eclipse is deep enough to be seen by eye in the diffuse light of 98~Tau using this interactive display, and is not associated with any of the background stars.

@@

% halo method
There are directions for improvement of the halo method itself, and for applying it beyond \kepler/\ktwo and TESS. It remains to be seen how well the method of optimizing convex objective functions can deal with significantly varying PSFs, such as from ground-based observations. The rapidly varying and moving seeing-limited PSF couples to flat field errors as is the case with \kepler, and leads to severe short-timescale instrumental noise. Self-calibration by the halo method, or a similar method, may permit improvements in ground-based photometry. Likewise, there may be other convex objective functions, including linear combinations of currently-used objective functions, which offer superior performance, for example by using combinations of different lagged functions to suppress systematics occurring at different timescales. The remaining unexplored space of convex objective functions may offer significant improvements on existing self-calibration techniques in high-cadence photometry and related problems in astronomy.
